---
title: "Hypothesis Test"
author: "Ying Jin"
date: "2022-11-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(ggplot2)
library(survminer)
library(rms)
library(survMisc)
```

Prepare the data

```{r}
pbc_data = 
  read_csv("./cirrhosis.csv") %>% 
  janitor::clean_names() %>%
  mutate(status = as.factor(status),
         drug = as.factor(drug),
         sex = as.factor(sex),
         ascites = as.factor(ascites),
         hepatomegaly = as.factor(hepatomegaly),
         spiders = as.factor(spiders),
         edema = as.factor(edema),
         stage = as.factor(stage)) %>% 
  mutate(status=recode(status,"D" =1, "C"=0,"CL"=0))
```

## Test for Treatment

### Log Rank Test

Using the log rank test, we want to test whether the occurrence of death in the D-penicillamine group ($S_1(t)$) delayed compared to the placebo group($S_0(t)$). The hypotheses are as below:$$H_0:S_1(t) = S_0(t)\\H_1:S_1(t) > S_0(t)$$

Conducting log rank test:

```{r}
# get rid of records which have no drug info

pbc_data_1 = 
  pbc_data %>% 
  drop_na(drug)

# K-M methods
pbc_kmsurvfit_1 = 
  survfit(formula = Surv(n_days,status)~drug,data = pbc_data_1)

# plot K_m curve
ggsurvplot(pbc_kmsurvfit_1,
           pval = TRUE, conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           risk.table.col = "strata", # Change risk table color by groups
           linetype = "strata", # Change line type by groups
           surv.median.line = "hv", # Specify median survival
           ggtheme = theme_bw(), # Change ggplot2 theme
           palette = c("#E7B800", "#2E9FDF"))

# log rank test
survdiff(Surv(n_days,status)~drug,data = pbc_data_1)
```

Since the $P\ value$ is $0.7 > 0.05$ (significance level), we fail to reject the null hypothesis and conclude that the survival probability in D-penicillamine group is not significantly different from that in the placebo group.

### Other Test

Using Gehan-Breslow generalized Wilcoxon, Tarone-Ware, Peto-Peto's, Fleming-Harrington test, the results are as follow:

```{r}
comp(ten(pbc_kmsurvfit_1))

```

We can see that using different weights, the test results are the same: the survival probability in D-penicillamine group is not significantly different from that in the placebo group.

These tests is unlikely to detect a difference when survival curves cross, which is the case in our data.

## Test for Stage

### Log Rank Test

Using the log rank test, we want to test whether the occurrence of death in different disease stage ($S_1(t)$， $S_2(t)$，$S_3(t)$ and $S_4(t)$ correspondingly) are different. The hypotheses are as below:$$H_0:S_1(t) = S_2(t)= S_3(t)= S_4(t)\\H_1:S_1(t)，S_2(t)，S_3(t) and S_4(t)$$ are not all equal.

Conducting log rank test:

```{r}
# get rid of records which have no stage info

pbc_data_2 = 
  pbc_data %>% 
  drop_na(stage)

# K-M methods
pbc_kmsurvfit_2 = 
  survfit(formula = Surv(n_days,status)~stage,data = pbc_data_2)

# plot K_m curve
ggsurvplot(pbc_kmsurvfit_2,
           pval = TRUE, conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           risk.table.col = "strata", # Change risk table color by groups
           linetype = "strata", # Change line type by groups
           surv.median.line = "hv", # Specify median survival
           ggtheme = theme_bw()) # Change ggplot2 theme
           #,palette = c("#E7B800", "#2E9FDF"))

# log rank test
survdiff(Surv(n_days,status)~stage,data = pbc_data_2)
```

Since the $P\ value$ is $4\times10^{-15} < 0.05$ (significance level), we reject the null hypothesis and conclude that at least two of the survival probability from the four disease stage groups is significantly different.

### Other Test

Using Gehan-Breslow generalized Wilcoxon, Tarone-Ware, Peto-Peto's, Fleming-Harrington test, the results are as follow:

```{r}
comp(ten(pbc_kmsurvfit_2))

```

We can see that using different weights, the test results are consistent with that of the log rank test.

Then we conduct multiple pairwise comparisons to see which groups have different survival probability.

```{r}
res_2 = pairwise_survdiff(Surv(n_days,status)~stage,data = pbc_data_2, p.adjust.method = "bonferroni",rho = 0)

res_2

symnum(res_2$p.value, cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
   symbols = c("****", "***", "**", "*", "+", " "),
   abbr.colnames = FALSE, na = "")

```

Base on above results， we can tell that the survival chance for patients in stage 4 is significantly lower than that of other patients, which can be seen clearly in the KM curve. 

## Test for Edema

### Log Rank Test

Using the log rank test, we want to test whether the occurrence of death for patients with different edema status ($S_N(t)$， $S_S(t)$ and $S_Y(t)$ correspondingly) are different. The hypotheses are as below:$$H_0:S_N(t) = S_S(t)= S_Y(t)\\H_1:S_N(t)，S_S(t)，and S_Y(t)$$ are not all equal.

Conducting log rank test:

```{r}
# get rid of records which have no edema info

pbc_data_3 = 
  pbc_data %>% 
  drop_na(edema)

# K-M methods
pbc_kmsurvfit_3 = 
  survfit(formula = Surv(n_days,status)~edema,data = pbc_data_3)

# plot K_m curve
ggsurvplot(pbc_kmsurvfit_3,
           pval = TRUE, conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           risk.table.col = "strata", # Change risk table color by groups
           linetype = "strata", # Change line type by groups
           surv.median.line = "hv", # Specify median survival
           ggtheme = theme_bw()) # Change ggplot2 theme
           #,palette = c("#E7B800", "#2E9FDF"))

# log rank test
survdiff(Surv(n_days,status)~edema,data = pbc_data_3)
```

Since the $P\ value$ is $2\times10^{-16} < 0.05$ (significance level), we reject the null hypothesis and conclude that at least two of the survival probability from the three edema groups is significantly different.

### Other Test

Using Gehan-Breslow generalized Wilcoxon, Tarone-Ware, Peto-Peto's, Fleming-Harrington test, the results are as follow:

```{r}
comp(ten(pbc_kmsurvfit_3))

```

We can see that using different weights, the test results are consistent with that of the log rank test.

Then we conduct multiple pairwise comparisons to see which groups have different survival probability.

```{r}
res_3 = pairwise_survdiff(Surv(n_days,status)~edema,data = pbc_data_3, p.adjust.method = "bonferroni",rho = 0)
res_3

symnum(res_3$p.value, cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, 0.1, 1),
   symbols = c("****", "***", "**", "*", "+", " "),
   abbr.colnames = FALSE, na = "")

```

Base on above results， we can tell that the survival chance for patients in the three edema groups is all different. In addition, according to the KM plot，we can see the trend is: $S_N(t)>S_S(t)>S_Y(t)$
