---
title: "Hypothesis Test"
author: "Ying Jin"
date: "2022-11-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(survival)
library(ggplot2)
library(survminer)
library(rms)
library(survMisc)
```

Prepare the data

```{r}
pbc_data = 
  read_csv("./cirrhosis.csv") %>% 
  janitor::clean_names() %>%
  mutate(status = as.factor(status),
         drug = as.factor(drug),
         sex = as.factor(sex),
         ascites = as.factor(ascites),
         hepatomegaly = as.factor(hepatomegaly),
         spiders = as.factor(spiders),
         edema = as.factor(edema),
         stage = as.factor(stage)) %>% 
  mutate(status=recode(status,"D" =1, "C"=0,"CL"=0))
```

## Log Rank Test

Using the log rank test, we want to test whether the occurrence of death in the D-penicillamine group ($S_1(t)$) delayed compared to the placebo group($S_0(t)$). The hypothese are as below:$$H_0:S_1(t) \le S_0(t)\\H_1:S_1(t) > S_0(t)$$

Conducting log rank test:

```{r}
# get rid of records which have no drug info

pbc_data_1 = 
  pbc_data %>% 
  drop_na(drug)

# K-M methods
pbc_kmsurvfit = 
  survfit(formula = Surv(n_days,status)~drug,data = pbc_data_1)

# plot K_m curve
ggsurvplot(pbc_kmsurvfit,
           pval = TRUE, conf.int = TRUE,
           risk.table = TRUE, # Add risk table
           risk.table.col = "strata", # Change risk table color by groups
           linetype = "strata", # Change line type by groups
           surv.median.line = "hv", # Specify median survival
           ggtheme = theme_bw(), # Change ggplot2 theme
           palette = c("#E7B800", "#2E9FDF"))

# log rank test
survdiff(Surv(n_days,status)~drug,data = pbc_data_1)
```

Since the $P\ value$ is $0.7 > 0.05$ (significance level), we fail to reject the null hypothesis and conclude that the survival probability in D-penicillamine group is significantly lower than that in the placebo group.

## Other Test

Using Gehan-Breslow generalized Wilcoxon, Tarone-Ware, Peto-Peto's, Fleming-Harrington test, the results are as follow:

Conducting log rank test:

```{r}
comp(ten(pbc_kmsurvfit))

```

We can see that using different weights, the test results are the same: the survival probability in D-penicillamine group is significantly lower than that in the placebo group.
