---
title: "KM estimator"
author: "Yiming Li"
output: html_document
---

```{r}
library(survival)
library(biostat3)
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(patchwork)
library(flexsurv)
library(MASS)
library(corrplot)
library(eha)
library(survminer)
```


```{r}
data = read.csv("./cirrhosis.csv") %>% 
  mutate(Status = case_when(
    Status == "CL" ~ 0,
    Status == "C" ~ 0,
    Status == "D" ~ 1,
  ))
data = na.omit(data)
data = data[,2:20]
```

## lifetable

```{r}
full = lifetab2(Surv(N_Days, Status == 1) ~ 1, data, breaks = seq(0, 5000, 500))
knitr::kable(full)
```

```{r}
hazard = data.frame(cbind(full[1:10,]$tstart, full[1:10,]$hazard, rep("full"))) %>% 
  rename(full = X3, hazard = X2, time = X1) %>% 
  mutate(time = as.numeric(time),
         hazard = as.numeric(hazard))

ggplot(hazard, aes(x = time, y = hazard)) +
  geom_point() + 
  geom_line() +
  labs(title = "Hazard function",
       x = "Time",
       y = "Hazard function")
```

```{r}
max(pull(data %>% filter(Sex == "M"), N_Days))
sexm = lifetab2(Surv(N_Days, Status == 1) ~ 1, data %>% filter(Sex == "M"), breaks = seq(0, 5000, 500))
sexm
```

```{r}
max(pull(data %>% filter(Sex == "F"), N_Days))
sexf = lifetab2(Surv(N_Days, Status == 1) ~ 1, data %>% filter(Sex == "F"), breaks = seq(0, 5000, 500))
sexf
```

based on drug
```{r}
max(pull(data %>% filter(Drug == "D-penicillamine"), N_Days))
drug = lifetab2(Surv(N_Days, Status == 1) ~ 1, data %>% filter(Drug == "D-penicillamine"), breaks = seq(0, 5000, 500))
drug
```

```{r}
max(pull(data %>% filter(Drug == "Placebo"), N_Days))
placebo = lifetab2(Surv(N_Days, Status == 1) ~ 1, data %>% filter(Drug == "Placebo"), breaks = seq(0, 5000, 500))
placebo
```

## Plot hazard function by trt on life-table estimate
```{r}
hazard = data.frame(rbind(cbind(drug[1:10,]$tstart, drug[1:10,]$hazard, rep("D-penicillamine")),
               cbind(placebo[1:10,]$tstart, placebo[1:10,]$hazard, rep("Placebo")))) %>% 
  rename(drug = X3, hazard = X2, time = X1) %>% 
  mutate(time = as.numeric(time),
         hazard = as.numeric(hazard))

ggplot(hazard, aes(x = time, y = hazard, color = drug)) +
  geom_point() + 
  geom_line() +
  labs(title = "Hazard Function Different Drug",
       x = "Time",
       y = "Hazard function")
```

## Plot hazard function by rxbased on life-table estimate
```{r}
hazard = data.frame(rbind(cbind(sexm[1:9,]$tstart, sexm[1:9,]$hazard, rep("sex m")),
               cbind(sexf[1:10,]$tstart, sexf[1:10,]$hazard, rep("sex f")))) %>% 
  rename(sex = X3, hazard = X2, time = X1) %>% 
  mutate(time = as.numeric(time),
         hazard = as.numeric(hazard))

ggplot(hazard, aes(x = time, y = hazard, color = sex)) +
  geom_point() + 
  geom_line() +
  labs(title = "Hazard function different sex",
       x = "Time",
       y = "Hazard function")
```


## Plot K-M survival function by sex
```{r}
km <- survfit(Surv(N_Days, Status) ~ Drug, data = data, type = c("kaplan-meier"))
autoplot(km) + 
  labs(x = "Survival time", y = "Survival prpbaibility",
  title = "KM Survival Time of different drug groups" )
```

## What is the median survival time for each sex group?
```{r}
summary(km)
```


## Compare survival function estimations between K-M and F-H methods
```{r}
fh <- survfit(Surv(N_Days, Status) ~ Drug, data = data, type = c("fleming-harrington"))
autoplot(fh) + 
  labs(x = "Survival time", y = "Survival prpbaibility",
  title = "FH Survival Time of different drug groups" )
```

```{r}
summary(fh)
```


## PH model

Select by significant level
```{r}
fit1 = coxph(Surv(N_Days, Status) ~ ., data = data)
summary(fit1)
```
Age, Edema, Bilirubin, Albumin, Copper, Prothrombin, Stage

## A submodel
```{r}
fit2 = coxph(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage, data = data)
summary(fit2)
```

```{r}
anova(fit1, fit2, test = 'F')
```

fail to reject null and conclude no difference


## backward selection
```{r}
step.model.back = step(fit1, direction = "backward")
```

Edema, Prothrombin, SGOT, Albumin, Copper, Age, Stage, Bilirubin



## A submodel based on backward
```{r}
fit3 = coxph(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data)
summary(fit3)
```

```{r}
anova(fit1, fit3, test = 'F')
```

## both selection
```{r}
step.model.both = step(fit1, direction = "both")
```

Cholesterol, Sex, Drug, Platelets, Hepatomegaly, Spiders, Alk_Phos, Tryglicerides, Ascites, Edema, Prothrombin, SGOT, Copper, Age, Stage, Bilirubin



## A submodel
```{r}
fit4 = coxph(Surv(N_Days, Status) ~ Cholesterol + Sex + Drug + Platelets + Hepatomegaly + Spiders + Alk_Phos + Tryglicerides + Ascites + Edema + Prothrombin + SGOT + Copper + Age + Stage + Bilirubin, data = data)
summary(fit4)
```

```{r}
anova(fit4, fit1, test = 'F')
```

## 
```{r}
surv_diff <- survdiff(Surv(N_Days, Status) ~ Sex, data = data)
surv_diff
```


## correlation plot
```{r}
data2 = data.frame(lapply(lapply(data,as.factor), as.numeric))
corrplot(cor(data2))
```

## Model Checking 
```{r}
#fit3 = coxph(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data)
par(mfrow = c(2, 4))
plot(cox.zph(fit3))
```

## parametric 
```{r}
fit_ph1 <- phreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "weibull")
summary(fit_ph1)
```


```{r}
fit_cox1 = coxreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2)
## same to coxph
summary(fit_cox1)  
```



```{r}
#(weibull", "gompertz", "lognormal", "llogis", "gengamma")
fit_ph2 <- phreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "gompertz")

fit_ph3 <- phreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "lognormal")

fit_ph4 <- phreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "loglogistic")
```

```{r}
par(mfrow = c(2, 2))
check.dist(fit_ph1, fit_cox1)
check.dist(fit_ph2, fit_cox1)
check.dist(fit_ph3, fit_cox1)
check.dist(fit_ph4, fit_cox1)
```

## the closest one
```{r}
fit_ph2 <- phreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "gompertz")

fit_cox1 = coxreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2)

check.dist(fit_ph2, fit_cox1)
```

## AFT and Parametric PH with gompertz
```{r}
fit_ph2 <- phreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "gompertz")
fit_aft1 <- aftreg(Surv(N_Days, Status) ~ Age + Edema + Bilirubin + Albumin + Copper + Prothrombin + Stage + SGOT, data = data2, dist = "gompertz")

max(fit_ph2$loglik)
max(fit_aft1$loglik)


```

